\section{The Privacy Externality Game}
\label{sec:game}

We formalise the strategic situation faced by data contributors in
sampling-based iDP as a normal-form game on a continuous action space. The
formalisation is a direct extension of \citet{pejo2019together}'s two-player
Collaborative-Learning game, with the central modification that the
\emph{privacy} cost is no longer separable across players.

\subsection{Players, actions and mechanism}

\begin{definition}[Privacy Externality Game]
\label{def:peg}
Fix a common $\delta>0$, an expected batch size $E_b$, an iteration count $I$,
and a partition of the dataset into $n$ \emph{groups} of sizes
$|G_1|,\dots,|G_n|$. The Privacy Externality Game (PEG) is the tuple
$\langle N, \Sigma, U\rangle$ where:
\begin{itemize}
\item $N = \{1,\dots,n\}$ is the set of players. We adopt the FL-silo
  granularity of \cite{kaiser2026your}: player $i$ is the entity controlling
  group $i$ (a silo). For a single-silo single-budget formulation each
  player picks a scalar $\eps_i$; the framework extends to per-record
  distributions inside a silo, which we discuss in
  \cref{sec:discussion}.
\item Each player's action set is $\Sigma_i = [\eps^{\min}, \eps^{\max}]
  \subset \R_{>0}$.
\item Given the action profile $\boldsymbol{\eps} \in \prod_i
  \Sigma_i$, the mechanism solution
  $(\sigma(\boldsymbol{\eps}), p_1(\boldsymbol{\eps}),\dots,
  p_n(\boldsymbol{\eps}))$ is determined by Definition~\ref{def:sampling-idp}.
\item Each player's utility is
  \begin{equation}
  u_i(\boldsymbol{\eps}) \;=\;
  B_i\cdot V(\sigma(\boldsymbol{\eps}))
  \;-\;
  C_i\cdot A_i(\boldsymbol{\eps}),
  \label{eq:utility}
  \end{equation}
  with $B_i, C_i > 0$.
\end{itemize}
\end{definition}

In the numerical evaluation we discretise $[\eps^{\min},\eps^{\max}]$ on a
log-spaced grid of $K=22$ points; we verify our equilibrium results are
stable to refinement.

\subsection{The two ingredients of utility}

\paragraph{Model-utility value $V(\sigma)$.}
We model model utility as a monotone-decreasing scalar function of the
mechanism's noise multiplier $\sigma$. We use
\begin{equation}
V(\sigma) \;=\; \frac{1}{1 + \sigma^2}. \label{eq:V}
\end{equation}
\emph{Frame.} The convex-DP-ERM excess-empirical-risk bound of
\citet{bassily2014private} (FOCS 2014) is, for a Lipschitz strongly-convex
loss, $O(d \log(1/\delta)/(n\eps)^2)$, which translates under the moments
accountant of \citet{abadi2016deep} (CCS 2016) into a $\sigma^2/n^2$
scaling of the excess risk. \emph{Validity in our setting.} The bound is
proven in the convex setting; deep-learning DP-SGD is generally non-convex
and the bound does not apply directly. We use $V(\sigma) = 1/(1+\sigma^2)$
as a \emph{directional} proxy. The empirical literature on DP-SGD in deep
learning~\cite{tramer2021private, abadi2016deep, boenisch2023have} reports
monotonic degradation of test accuracy with $\sigma$, supporting the
qualitative form. None of our game-theoretic results depend on the precise
shape of $V$: only that it is continuous, monotone-decreasing in $\sigma$,
and bounded.

\paragraph{Realised MIA advantage $A_i(\boldsymbol{\eps})$.}
We use the tight analytical MIA advantage of the $I$-fold $(p_i, \sigma)$-
subsampled Gaussian as defined in~\eqref{eq:adv}. This is the same advantage
measure used by \citet{kaiser2026your}; our choice transfers their
empirical measurement to a strategic setting unchanged. The functional
dependence $A_i(\boldsymbol{\eps})$ runs through $(p_i, \sigma)$, both of
which are determined by the entire vector $\boldsymbol{\eps}$ via the
mechanism's (C1) and (C2).

\subsection{Decomposition into contracted and excess cost}
\label{sec:decomposition}

The privacy cost $A_i(\boldsymbol{\eps})$ in~\eqref{eq:utility} can be
decomposed into a contracted (separable) part and an excess (coupled) part
that makes the externality explicit. Let $\bar c(\eps_i)$ denote the
MIA advantage of a reference, non-interdependent $(\eps_i, \delta)$-DP
mechanism (e.g., the post-processing-tight $(\eps_i,\delta)$-DP advantage
upper bound $(e^{\eps_i}-1+2\delta)/(e^{\eps_i}+1)$ of
\cite{hayes2024bullseye, kaissis2024beyond}). The realised cost is
\begin{equation}
A_i(\boldsymbol{\eps}) \;=\; \bar c(\eps_i) - \mathrm{slack}_i(\boldsymbol{\eps}),
\label{eq:decomp}
\end{equation}
where the \emph{slack} $\mathrm{slack}_i(\boldsymbol{\eps}) := \bar c(\eps_i)
- A_i(\boldsymbol{\eps}) \geq 0$ is the amount by which the realised
advantage falls below the worst-case-bound at $(\eps_i, \delta)$. In
sampling-based iDP this slack is a function of the entire $\boldsymbol{\eps}$
through the mechanism's coupling. \emph{The externality reduces the slack:}
when others lower their $\eps_j$, $\mathrm{slack}_i$ shrinks (player $i$
moves closer to its $(\eps_i, \delta)$-DP worst case). This is the formal
quantity that the proposed $(\eps_i, \delta, \Delta)$-iDP
contract~\cite{kaiser2026your, kaissis2024beyond} would cap: it would
require $\mathrm{slack}_i$ to differ from a baseline reference by at most
$\Delta$. In this sense, \emph{the $\Delta$-bound is exactly the device that
neutralises the externality}.

We will work directly with $A_i$ in the utility~\eqref{eq:utility} for
analytical simplicity; the decomposition matters as a conceptual bridge to
PoP~\cite{pejo2019together} (whose cost $c(p_i)$ was separable, hence
$\mathrm{slack}_i$ did not exist) and to the iDP literature
\cite{kaiser2026your} (where $\mathrm{slack}_i$ \emph{is} the new
externality term).

\subsection{Vulnerability function and feasibility}

\begin{definition}[Vulnerability function]
\label{def:vulnerability}
The \emph{vulnerability function} of player $i$ is $A_i\colon
\prod_i\Sigma_i \to [0, 1]$, $\boldsymbol{\eps} \mapsto
\Adv\bigl(\mathcal{M}_{p_i(\boldsymbol{\eps}),\,\sigma(\boldsymbol{\eps}),\,I}\bigr)$.
\end{definition}

\begin{assumption}[Participation]
\label{ass:participation}
A profile $\boldsymbol{\eps}$ is \emph{feasible} if and only if the mechanism
solution satisfies $p_i(\boldsymbol{\eps}) > 0$ for every player $i$.
Profiles with $p_i = 0$ correspond to a player whose data never enters
training; their iDP contract is satisfied trivially. We exclude such
profiles from the game; equivalently, the action space $\Sigma_i$ is
implicitly restricted at each move to the participating region.
\end{assumption}

\subsection{Coalition variant}

\begin{definition}[$k$-coalition Stackelberg PEG]
\label{def:coalition}
Fix a target index $t \in N$ and a colluding subset $C \subset
N\setminus\{t\}$ with $|C| = k$. The $k$-coalition Stackelberg PEG is the
two-stage game in which:
\begin{enumerate}
\item Stage~1 (leader): the coalition jointly commits to an action vector
$\boldsymbol{\eps}_C$ to maximise the augmented coalition utility
\begin{equation}
\sum_{i \in C} u_i(\boldsymbol{\eps}) \;+\;
\lambda\, A_t(\boldsymbol{\eps}),
\label{eq:coalition-util}
\end{equation}
where $\lambda \geq 0$ is the \emph{malice weight} on the target's
vulnerability~\cite{pejo2019together}. $\lambda = 0$ recovers the benign
PEG; $\lambda \to \infty$ recovers the
budget-manipulation attack of \citet{kaiser2026your}.
\item Stage~2 (followers): the non-coalition players $\{t\} \cup (N
\setminus (C \cup \{t\}))$ best-respond.
\end{enumerate}
\end{definition}

\subsection{Equilibrium concepts}

We use the standard Nash equilibrium concept: an $\boldsymbol{\eps}^*$ in
the feasible region such that $u_i(\eps_i^*, \boldsymbol{\eps}_{-i}^*) \geq
u_i(\eps_i, \boldsymbol{\eps}_{-i}^*)$ for every player $i$ and every
feasible deviation $\eps_i$. Existence of pure NE on the continuous action
space is not automatic; we report pure NE found by brute search on a fine
log-spaced grid and verify the result is stable under grid refinement.
Where no pure NE exists on the dense grid we report the empirical mixed
strategy reached by fictitious play~\cite{monderer1996potential}.

\begin{remark}[Awareness as information, not utility]
\label{rem:awareness}
A useful distinction is between players who can \emph{compute}
$A_i(\boldsymbol{\eps})$ given the public mechanism's parameters and players
who can only access their nominal $(\eps_i,\delta)$-iDP contract. Both
populations face the same utility~\eqref{eq:utility}; they differ only in
the rationality they apply. A \emph{nominal-information} (``naive'') player
who treats the contracted bound $\bar c(\eps_i)$ as a separable cost plays
the PoP game of \citet{pejo2019together} unchanged. An
\emph{aware} player who computes $A_i(\boldsymbol{\eps})$ from the
mechanism's parameters plays the PEG. We will return to this distinction in
\cref{sec:discussion}.
\end{remark}
