\section{Discussion and Conclusion}
\label{sec:discussion}

We have shown that sampling-based individual differential privacy is not
just \emph{vulnerable} to collusion attacks but \emph{equilibrium-asymmetric}.
The pure Nash equilibria of the Privacy Externality Game, when they exist on
the dense action grid, sit \emph{off the diagonal}: one player accepts a
higher realised MIA advantage and is compensated by a higher model-utility
benefit; the other absorbs the opposite trade. The budget-manipulation
attack of \citet{kaiser2026your} is precisely the asymmetric equilibrium
played by a coalition that weights the target's harm in its joint utility,
and \cref{prop:coalition-attack} shows this is the rational best response of
\emph{any} coordinated subset of data holders.

\paragraph{Limitations of this work.} Three caveats are worth noting.
First, our model assumes one-shot full-information play. Real
deployments of iDP are dynamic: contributors join and leave; budgets are
re-negotiated; trainers re-train. A natural extension is a repeated game
in which the players also have private preference types and learn about
one another across rounds. \cref{app:bayesian} sketches such a game ---
the \emph{Bayesian Repeated PEG} --- as a direct instance of the
Stackelberg-learning template of \citet{letchford2009learning,
blum2014learning, balcan2015commitment} applied to the iDP mechanism,
together with a small simulation. We treat it as a brief extension here
and do not pursue regret bounds or the fully two-sided repeated game
of \citet{aumann1995repeated}.
Second, we assume full information about other players' actions. Real
contracts may keep budgets private, and \cref{rem:awareness} already
distinguishes between fully aware players and players who only see their
nominal $(\eps_i,\delta)$ contract. A Bayesian PEG that treats $\eps_i$ as
a private type drawn from a known distribution is a natural and important
extension.
Third, our analysis is restricted to the sampling-based realisation of iDP.
The sensitivity-based realisation~\cite{boenisch2023have} adjusts per-sample
clipping bounds rather than per-sample sampling rates and does not exhibit
the same externality (the noise-to-sensitivity ratio is preserved by
construction). Our framework specialises to that case as the trivial PEG in
which the externality vanishes.

\paragraph{The role of the model-utility function $V$.} Our choice $V(\sigma)
= 1/(1+\sigma^2)$ is motivated by the convex DP-ERM excess-risk bound of
\citet{bassily2014private}, and we use it as a directional proxy in our
possibly non-convex setting. The qualitative findings---existence of
asymmetric NEs, monotone coalition attack, externality lemma---do not depend
on the precise shape of $V$; they require only that $V$ is continuous,
monotone-decreasing in $\sigma$, and bounded. We expect the same qualitative
picture under any utility function with these three properties.

\paragraph{Mitigation.} The natural mitigation is to make the contract
internalise the externality. The $(\eps_i, \delta, \Delta)$-iDP formulation
of \citet{kaissis2024beyond} adds a $\Delta$-divergence bound on the realised
vulnerability, restricting the action space of the PEG to profiles that
satisfy a per-record bound on excess advantage. We expect this restriction
to admit a unique equilibrium and to drive the realised advantage gap
between players to within $\Delta$ by definition. A clean
characterisation of the constrained PEG is left to future work.

\paragraph{Open questions.} The following questions naturally arise from
this work but lie outside its scope:
\begin{enumerate}
\item Does a pure-strategy NE always survive on the continuous action space
under the coupled cost, and does a $\Delta$-bound on the slack suffice for
existence?
\item Is the PEG a potential game in the sense of \cite{monderer1996potential}?
PoP's analysis was a potential game because the privacy cost was
separable; the externality term breaks this directly. Conditions on the
cross-partial of $A_i$ that recover the potential structure would unify
the two settings.
\item What is the sign of the externality across composition regimes
(minority vs majority weak-privacy)? Our experiments suggest the smaller
group is favoured, but the analytic dependence on $|G_g|$ deserves a
closed-form characterisation.
\item What is the appropriate player granularity (single sample versus
silo) in the FL setting, and how does the granularity change the
equilibria?
\item Worst-case (analytical) vs realised (empirical) cost---which should
the contract be written against, and how large is the gap in practice
across deep-learning workloads?
\item Is the $(\eps_i, \delta, \Delta)$-iDP contract by itself sufficient
as a collusion-resistant mechanism, or are sensitivity-based iDP or
public-data padding needed when the slack is exceeded?
\end{enumerate}

\paragraph{Wider lesson.} The wider lesson is that an \emph{individual}
privacy contract is incomplete: it does not pin down the collective
dynamics. Where individuals interact strategically (in a single
machine-learning training, in a federated round, in any future composition
of the iDP mechanism), the contract must also constrain their joint
behaviour or the equilibrium will quietly violate at least one
participant's expectations while preserving every formal guarantee.

\paragraph{Conclusion.} The promise of iDP---that every data contributor
controls her own privacy---is broken by the privacy externality of
sampling-based mechanisms. We formalised the strategic situation faced by
iDP data contributors as the Privacy Externality Game, proved monotonicity
of the externality and characterised the equilibria. The result is that
the collusion attack on iDP is not adversarial but \emph{rational}, and the
remedy must operate at the level of the contract.
