\appendices

\section{Detailed mechanism-solver construction}
\label{app:mechanism}

We elaborate on the bisection used to obtain the mechanism solution
$(\sigma, p_1,\dots,p_n)$ for an action profile $\boldsymbol{\eps}$.

\paragraph{Inner step: $p_g$ from $(\sigma, \eps_g)$.}
For each fixed $\sigma$ and each group~$g$, we want the largest $p_g \in [0,1]$
satisfying the $I$-fold $(p_g, \sigma)$-subsampled Gaussian at $(\eps_g,
\delta)$-DP under (C1). Since the $\eps$ value of the subsampled Gaussian is
strictly increasing in $p_g$ (more sampling $\Rightarrow$ more privacy
loss), a unique $p_g \in (0, 1]$ exists, capped at $1$ if the budget cannot
be exhausted. We compute the per-group $\eps$ by R\'enyi-DP
composition~\cite{mironov2017renyi} and binary-search $p_g$ to bring this
$\eps$ to $\eps_g$ to within a relative tolerance of $10^{-3}$.

\paragraph{Outer step: $\sigma$ from $(p_g, \boldsymbol{\eps})$.}
Define $B(\sigma) = \sum_g |G_g|\, p_g(\sigma, \boldsymbol{\eps})$ as the
realised expected batch size. Since each $p_g$ is monotone-increasing in
$\sigma$, so is $B$. We bisect $\sigma$ to bring $B$ to the target $E_b$,
again to within a relative tolerance of $10^{-3}$.

\paragraph{Participation threshold.}
We declare profile infeasible (per \cref{ass:participation}) if any
$p_g$ falls below $10^{-5}$. Below this threshold the group's contribution
to (C2) is numerically negligible and the iDP contract is trivially
satisfied. We verified the qualitative results are insensitive to the
choice of this threshold across $[10^{-6}, 10^{-3}]$.

\paragraph{MIA advantage.}
Given the mechanism solution $(\sigma, p_g)$, we compute the MIA advantage
of group~$g$ exactly from the privacy-loss-distribution of the $I$-fold
$(p_g, \sigma)$-subsampled Gaussian using
\cite{google-dp-accounting}, then evaluate the trade-off function and
\cref{eq:adv}. This is the same formulation used by
\citet{kaiser2026your} and yields the tightest analytical MIA advantage
for the resulting mechanism.

\section{Validity of $V(\sigma) = 1/(1+\sigma^2)$}
\label{app:V}

\paragraph{Frame.} The convex-DP-ERM excess-risk bound of
\citet{bassily2014private} is, for a 1-Lipschitz $L$-strongly-convex
$\ell\colon \mathcal{X} \times \mathcal{D} \to \R$,
\[
\E[\hat L(\widehat\theta) - \hat L(\theta^*)] = O\Bigl(\frac{d \log(1/\delta)}{n^2 \eps^2}\Bigr).
\]
Translating to noise multiplier via the moments accountant of
\citet{abadi2016deep}, this reads $O(\sigma^2 \cdot d / n^2)$ in
\emph{excess} empirical risk. Treating ``model utility'' as
$-$``excess risk'' and re-normalising to $V \in [0,1]$ produces $V \propto
1/(1+\alpha\sigma^2)$ for a problem-dependent $\alpha$. We set $\alpha = 1$
for simplicity; sensitivity analysis (\cref{app:sensitivity}) shows the
qualitative results are unchanged for $\alpha \in [0.1, 10]$.

\paragraph{Beyond convex.} For non-convex deep-learning losses no analogous
closed-form bound is known. The empirical
literature~\cite{tramer2021private, abadi2016deep, boenisch2023have}
reports monotone test-accuracy degradation with $\sigma$ in standard
benchmarks. We use $V(\sigma) = 1/(1+\sigma^2)$ as a \emph{directional}
proxy; the qualitative results of our paper (existence of asymmetric NEs,
monotone coalition attack, externality lemma) do not depend on the precise
shape of $V$; they require only that $V$ is continuous, monotone-decreasing
in $\sigma$, and bounded.

\section{Sensitivity to grid resolution and to $V$}
\label{app:sensitivity}

\paragraph{Grid refinement.} For the 2-player setting of \cref{fig:C1} we
re-ran the sweep with $K = 44$ log-spaced points (twice the resolution) and
verified that the set of pure NEs is contained, modulo the new
intermediate grid points, in the set found at $K = 22$. Specifically, the
asymmetric NEs at $K=22$ persist at $K=44$ and the new intermediate grid
points do not introduce additional NEs.

\paragraph{Functional form of $V$.} The same 2-player sweep was repeated
with $V(\sigma) = e^{-\sigma}$ (an exponential decay alternative); the pure
NEs shift quantitatively (mostly toward higher $\eps$ values, since
$V_{\exp}$ falls off faster than $V_{1/(1+\sigma^2)}$ as $\sigma$ grows)
but the asymmetric structure is preserved.

\paragraph{Participation threshold.} Lowering the threshold to $10^{-6}$
extends the feasible region; the NEs found at $10^{-5}$ remain NEs but
additional ones appear in the newly feasible interior. Raising the threshold
to $10^{-3}$ shrinks the feasible region and removes some NEs at the
boundary; the surviving NEs do not move.
