\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{array}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{subcaption}
\usepackage[hidelinks]{hyperref}
\usepackage[capitalise]{cleveref}

% Theorem environments
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{observation}[theorem]{Observation}

% Custom macros
\newcommand{\eps}{\varepsilon}
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Adv}{\mathrm{Adv}}
\newcommand{\NE}{\mathrm{NE}}
\newcommand{\BR}{\mathrm{BR}}
\newcommand{\PoEV}{\mathrm{PoEV}}
\newcommand{\PoA}{\mathrm{PoA}}
\newcommand{\PEG}{\mathrm{PEG}}
\newcommand{\iDP}{i\text{-}\mathrm{DP}}
\newcommand{\DP}{\mathrm{DP}}
\newcommand{\MIA}{\mathrm{MIA}}
\newcommand{\GS}[1]{|G_{#1}|}

% Title and authors
\title{The Privacy Externality Game:\\
Strategic Budget Choices in Sampling-Based Individual Differential Privacy}

\author{
\IEEEauthorblockN{Anonymous Submission}
}

\begin{document}
\maketitle

\begin{abstract}
Individual differential privacy (iDP) gives each data contributor a personal
privacy budget. The currently dominant realisation---sampling-based iDP---turns
each contributor's chosen budget into a per-sample inclusion rate, with a single
noise multiplier shared across the dataset. We show that this realisation creates
a privacy externality: lowering one's own budget mechanically raises the membership-
inference (MIA) advantage on every other data point. Although every individual
$(\eps_i, \delta)$-iDP guarantee is honoured, the realised privacy of any one
contributor is determined by the choices of all others. We model the resulting
strategic situation as a finite, normal-form game, the Privacy Externality Game
(PEG), and study its equilibria. We prove that the privacy externality is
monotone: $\partial A_i / \partial \eps_j \leq 0$ for $j \neq i$ under the
mechanism's binding constraints. We then characterise the game under two
realistic awareness regimes---naive (players perceive only their nominal
$\eps_i$) and aware (players perceive the realised MIA advantage). In a
canonical $2$-player setting we find that both regimes admit \emph{asymmetric}
pure Nash equilibria: the naive game equilibrates at $(\eps_i, \eps_j) = (1, 8)$
or $(8, 1)$ with one player exposed at $A = 0.275$ and the other at $A = 0$;
the aware game equilibrates at $(1, 2)$ or $(2, 1)$, internalising the
externality but leaving the asymmetry intact. The aware equilibrium has
roughly half the total realised MIA of the naive equilibrium, showing that
\emph{awareness mitigates but does not eliminate} the externality. A coalition
Stackelberg analysis recovers the attack of Kaiser et al.\ (SaTML 2026) as the
rational best response of any coordinated set of data holders. The implication is that the empirical
$62\%$-vulnerability attack rate observed in iDP is not adversarial but
equilibrium behaviour: even non-malicious players in iDP impose privacy
externalities on every other contributor.
\end{abstract}

\begin{IEEEkeywords}
Differential privacy, individual differential privacy, game theory, Nash
equilibrium, collusion, membership inference.
\end{IEEEkeywords}

\input{tex/01_intro}
\input{tex/02_background}
\input{tex/03_game}
\input{tex/04_analysis}
\input{tex/05_experiments}
\input{tex/06_related}
\input{tex/07_discussion}

\bibliographystyle{IEEEtran}
\bibliography{refs}

\input{tex/08_appendix}

\end{document}
